---
title: "Data Wranging in R - Intro to tidyr and dplyr"
author: "Greg Clunies"
date: "3/24/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install tidyr and dplyr
```{r}
# install.packages('tidyr')
library(tidyr)

# install.packages('dplyr')
library(dplyr)

# load in mtcars dataset (a built in R dataset)
data(mtcars)

# convert the data to a tbl class
table1 <- tbl_df(mtcars)

View(table1)  # view data in spreadsheet-like display (note capital V)
```

## Setup and Data for this Tutorial
```{r}
# install.packages("devtools")
# devtools::install_github("rstudio/EDAWR")
library(EDAWR)
?storms
?cases
?pollution
?tb
```


# tidyr

## Example 1 - gather()
```{r}
library(tidyr)

?gather
?spread

cases
gather(cases, key='year', value='n', '2011','2012','2013')
# OR
gather(cases, key = 'year', value = 'n',  2:4)  # this seems preferable if a lot of columns to be "gathered"
```

## Example 2 - spread()
```{r}
?pollution

pollution

# spread example
pollution2 <- spread(pollution, 'size', value = 'amount')
pollution2

# inverse of spread is gather!
gather(pollution2, key = 'size', value = 'amount', 2:3)
```
Notice how gather() and spread() are inverse operations!

## Example 3 - separate()
```{r}
storms

# notice that there are still some variables "hidden" in this dataset (year, month, day)

separate(storms, date, into = c("year", "month", "day"), sep = "-")
# now we have year, month, day as separate variables
```

## Example 3 - unite()
```{r}
storms

# notice that there are still some variables "hidden" in this dataset (year, month, day)

storms2 <- separate(storms, date, into = c("year", "month", "day"), sep = "-")
storms2
# now we have year, month, day as separate variables

unite(storms2, "date", year, month, day, sep='-')
```


# dplyr
---
```{r}
library(dplyr)
?select
?filter
?arrange
?mutate
?summarise
?group_by
```
**Ways to Access Infromation**

1. Extract existing variables -- select()
2. Extract existing observations -- filter()
3. Derive new variables (rom existing) -- mutate()
4. chnge unit of analysis -- summarise()

## Example 1 - select()
```{r}
storms

select(storms, storm, pressure)

select(storms, -storm) # drop 'storm' column

select(storms, wind:date) # select range of columns
```
**Useful select() functions**
- , Select everything but
: , Select range
contains() - Select columns whose name contains a character string
ends_with() - Select columns whose name ends with a string
everything() - Select every column
matches() - Select columns whose name matches a regular expression
num_range() - Select columns named x1, x2, x3, x4, x5
one_of() - Select columns whose names are in a group of names
starts_with() - Select columns whose name starts with a character string

## Example 2 - filter()
```{r}
storms

filter(storms, wind >= 50)

filter(storms, wind >= 50, storm %in% c("Alberto", "Alex", "Allison"))
```
**logical test in R**

*?Comparison*
< , Less than
> , Greater than
== , Equal to
<= , Less than or equal to
>= , Greater than or equal to
!= , Not equal to 
%in% , Group membership
is.na , Is NA
!is.na , Is not NA

*?base::Logic*
& , boolean and
| , boolean or
xor , exactly or
! , not
any , any true
all , all true

## Example 3 - mutate()
```{r}
storms

mutate(storms, ratio = pressure/wind)

mutate(storms, ratio = pressure/wind, inverse = ratio^-1)
```
**useful mutate() functions**
pmin(), pmax()  - Element-wise min and max
cummin(), cummax() - Cumulative min and max
cumsum(), cumprod() - Cumulative sum and product
between() - Are values between a and b?
cume_dist() - Cumulative distribution of values
cumall(), cumany() - Cumulative all and any
cummean() - Cumulative mean
lead(), lag() - Copy with values one position
ntile() - Bin vector into n buckets
dense_rank(), min_rank(), percent_rank(), row_number() - Various ranking methods

## Example 4 - summarise()
```{r}
summarise(pollution, mean = mean(amount), sum = sum(amount), n = n())

# using piping - think of it as: "pass 'pollution' as the first argument to the summarise function"
pollution %>% summarise(mean = mean(amount), sum = sum(amount), n = n())
```

## Example 5 - arrange()
```{r}
storms

arrange(storms, wind)

arrange(storms, desc(wind))

arrange(storms, wind)
arrange(storms, wind, date) # will arange (sort) by wind, THEN date
```

## Example 6 - The pipe operator, %>%
```{r}
select(storms, storm, pressure)  # THIS, is the same as.... 

storms %>% select(storm, pressure)  # THIS

# this pipe operator can help us "chain" operations together

storms %>%
  mutate(ratio = pressure / wind) %>%
  select(storm, ratio)
```

%>% Shortcut is: Cmd + Shift + M

## Example 7 - group_by()
```{r}
pollution %>% group_by(city)

pollution %>% 
  group_by(city) %>%
  summarise(mean = mean(amount), sum = sum(amount), n = n())

pollution 

pollution %>% ungroup()  # not sure what this line does

# here is teh answer
?ungroup
```

