---
title: 'Lesson_3: Explore One Variable'
author: "Greg Clunies"
date: "3/18/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set working directory
```{r}
# macbook path
setwd("/Users/GregoryClunies/Repos/Udacity/Exploratory_Data_Analysis/L3_Explore_One_Variable")
```


## Read in Data
```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

## Histogram of User's birthdays
```{r}
# install.packages('ggplot2')
library(ggplot2)

# install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)

theme_set(theme_minimal(16))

names(pf)
qplot(data = pf, x = dob_day) +
  scale_x_continuous(breaks = 1:31)

```

## Faceting
```{r}
qplot(data = pf, x = dob_day) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)
```

## Friend Count
```{r}
# plot the friend count data
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
qplot(data = pf, x = friend_count)

# this is the alternate method using ggplot() instead of qplot()
ggplot(data = pf, aes(x = friend_count)) + 
  geom_histogram()
```

## Limiting the axes
```{r}
qplot(data = pf, x = friend_count, xlim = c(0, 1000))

# aleternate solution 1
qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0, 1000))

# alternate solution 2
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```

## Adjusting bin Width
```{r}
ggplot(data = pf, aes(x = friend_count)) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0,1000,50)) +
  facet_wrap(~gender)
```

## Ommitting NA observations
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x = friend_count)) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0,1000,50)) +
  facet_wrap(~gender, ncol=2)
```

## Statistics 'by' gender
```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```


## Investiagting Tenure (how long a user has been on FB) - Using colors to make plots pretty
```{r}
ggplot(data = pf, aes(x = tenure/365)) +
  geom_histogram(binwidth = 0.5, color = 'black', fill = 'blue')
```

## Labelling Plots
```{r}
ggplot(data = pf, aes(x = tenure/365)) +
  geom_histogram(binwidth = 0.25, color = 'black', fill = 'blue') +
  scale_x_continuous(breaks = seq(1,7,1), limits =c(0,7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

## User Ages
```{r}
ggplot(data = pf, aes(x = age)) +
  geom_histogram(binwidth = 1, color = 'black', fill = 'orange') +
  scale_x_continuous(breaks = seq(0,113,5), limits = c(0,120)) +
  xlab('User Age') +
  ylab('Number of users in sample')
```

## Transforming Data
```{r}
# install.packages('gridExtra')
library(gridExtra)

# make the individual plots
p1 <- ggplot(data = pf, aes(x = friend_count)) +
    geom_histogram(color = 'black', fill = 'blue') 

p2 <- p1 + scale_x_log10()

p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

## Frequency Polygons
```{r}
# y-axis as counts
p1 <- ggplot(data = subset(pf, !is.na(gender)), aes(x = www_likes)) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10() +
  xlab('Likes') +
  ylab('Count of users')

# y-axis as proportions... need to add 'y' to aes
p2 <- ggplot(data = subset(pf, !is.na(gender)), aes(x = www_likes, y = ..count../sum(..count..))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10() +
  xlab('Likes') +
  ylab('Proportion of users')

grid.arrange(p1, p2)
```

## Which gender has more likes?
```{r}
by(pf$www_likes, pf$gender, sum)
```

## Box Plots
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x = gender, y = friend_count)) +
  geom_boxplot() + 
  scale_y_continuous(limits = c(0,1000))  # using scale_y_continuous() actually removes data BEFORE plotting the boxes

# instead we should use coord_cartesion leyer
ggplot(data = subset(pf, !is.na(gender)), aes(x = gender, y = friend_count)) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0,1000))  
# note that we did not get a warning about removing rows of data this time, like we did when using scale_y_continuous()
# the box plots will have slightly different shape as well since we have not ommitted any data
```

## Box Plots, Quartiles, and Friend Counts
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x = gender, y = friend_count)) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

## Box Plots, Quartiles, and Friend Requests
```{r}
ggplot(data = subset(pf, !is.na(gender)), aes(x = gender, y = friendships_initiated)) +
  geom_boxplot() + 
  coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

## Getting logical

We want to know how many users access facebook from the mobile app relative to the dektop app. This is a good starting point to understand where development focus should be.
```{r}
summary(pf$mobile_likes)

# use mobile likes as a proxy for users who use mobile app
summary(pf$mobile_likes > 0)

# create mobile_check_in variable
pf$mobile_check_in <- NA  # assign column of NA's to dataframe
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)  # convert to a factor
summary(pf$mobile_check_in)

p_mobile <- sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
p_mobile
```


