Lesson 5
========================================================

### Multivariate Data
Notes:

## setup and libraries
```{r}
setwd('/Users/GregoryClunies/Repos/Udacity/Exploratory_Data_Analysis/L7_Explore_Many_Variables')

library(tidyverse)

pf <- read.delim('pseudo_facebook.tsv')
```


```{r}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)
```


### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
pf.fc_by_age_gender <- pf %>% 
  filter(!is.na(gender)) %>% 
  group_by(age, gender) %>% 
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>% 
  ungroup() %>% 
  arrange(age)

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(data = pf.fc_by_age_gender, aes(x = age, y = median_friend_count)) +
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
# install.packages('reshape2')
#library(reshape2)

# DONT USE RESHAPE2, USE 'tidyr' package INSTEAD

# I have already imported the tidyr library earlier by importing the 'tidyverse' library

pf.fc_by_age_gender.wide <- spread(subset(pf.fc_by_age_gender,
              select = c('gender', 'age', 'median_friend_count')),
       gender,
       median_friend_count)

head(pf.fc_by_age_gender.wide)
```
To switch the data back to the long format, we use `gather()`.
Note, since we only have median friend count 
```{r}
pf.fc_by_age_gender.long <- gather(data = pf.fc_by_age_gender.wide,
                                   key = 'gender',
                                   value = 'median_friend_count',
                                   female,
                                   male
                                   ) %>% 
arrange(age)
head(pf.fc_by_age_gender.long )
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(data = pf.fc_by_age_gender.wide,
       aes(x = age, y = female/male)) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)
?cut

pf$year_joined.bucket <- cut(pf$year_joined,
                             c(2004,2009, 2011, 2012, 2014)
                             )
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket)

ggplot(data = subset(pf, !is.na(year_joined.bucket)),
       aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(data = subset(pf, !is.na(year_joined.bucket)),
       aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(data = pf.fc_by_age_gender, aes(x = age, y = median_friend_count), linetype=2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
library(zoo)

ggplot(data = subset(pf, tenure >= 1 ),
       aes(x = tenure, y = friendships_initiated/tenure)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set


### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
# str(yo)

# Change id from and int to a factor
yo$id <- factor(yo$id)
str(yo)

ggplot(data = yo, aes(x = price)) +
  geom_histogram()
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

summary(yo$all.purchases)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(data = yo, aes(x = time, y = price)) +
  geom_jitter(alpha = 1/10, height = 0.5)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)

sample.ids <- sample(levels(yo$id), 16)

ggplot(data = subset(yo, id %in% sample.ids),
       aes(x = time, y = price)) +
  geom_point(aes(size = all.purchases), shape = 1) + 
  geom_line() +
  facet_wrap(~id) 
```

# same analysis using a different seed (will sample 16 different household ids)
```{r}
set.seed(1234)

sample.ids <- sample(levels(yo$id), 16)

ggplot(data = subset(yo, id %in% sample.ids),
       aes(x = time, y = price)) +
  geom_point(aes(size = all.purchases), shape = 1) + 
  geom_line() +
  facet_wrap(~id) 
```


***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r}
# install.packages("GGally")
library(GGally)
theme_set(theme_minimal(20))

# set seed for reproducible results
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ], axisLabels = 'internal')


```



### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

