---
output: html_document
editor_options: 
  chunk_output_type: inline
---
 Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

```{r}
setwd('/Users/GregoryClunies/Repos/Udacity/Exploratory_Data_Analysis/L5_Explore_Two_Variables')

library(tidyverse)
```

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_point(alpha = 1/20) + 
  xlim(13, 90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:
Jitter lets us distinguish each point more easily by adding in some "noise" to help set points apart.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_point(alpha = 1/20) + 
  xlim(13, 90) + 
  coord_trans(y = 'sqrt')
```
**NOTE:**
geom_jitter will not work with coord_trans() ...need to look into this. Must use geom_point instead. See below code chunk to see how to add a "jitter" to geom_point() plots using the 'position = position_jitter()' argument (see below).

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

# Coords trans and jitter
```{r}
ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) + 
  coord_trans(y = 'sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(data = pf, aes(x = age, y = friendships_initiated)) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(tidyverse)

age_groups <- group_by(pf, age)
pf.fc_by_age<- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())

head(pf.fc_by_age)

# OR, we can do the same thing using the pipe operator, %>%, which allows you to 'chain' functions toegther, one after the other

pf.fc_by_age <- pf %>% 
  group_by(age) %>% 
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>% 
  arrange(age)

head(pf.fc_by_age, desc=True, 20)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(data = pf.fc_by_age, aes(x = age, y = friend_count_mean)) +
  geom_line() + 
  scale_x_continuous(breaks = seq(10,90,10), limits = c(10,90))
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_point(alpha = 1/20,
             position = position_jitter(h = 0),
             color = 'orange') + 
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary',
            fun.y = quantile,
            fun.args = list(probs = 0.1),
            linetype = 2,
            color = 'blue') +
  geom_line(stat = 'summary',
            fun.y = quantile,
            fun.args = list(probs = 0.5),
            color = 'blue') +
  geom_line(stat = 'summary',
            fun.y = quantile,
            fun.args = list(probs = 0.9),
            linetype = 2,
            color = 'blue')
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')

# ALTERNATE CODE

with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <=70), cor.test(age, friend_count, method = 'pearson'))
```
Genaeral rules for correlation coeff:
abs(r) > 0.3 ---> slight significance
abs(r) > 0.5 ---> moderate significance
abs(r) > 0.7 ---> significant

An acceptable value of significant may vary based on the field/application

***

### Correlation Methods
Notes:
Quick reference on correlation methods:
http://www.statisticssolutions.com/correlation-pearson-kendall-spearman/
```{r Correlation on Subsets}
# Pearson r
c1 <- with(subset(pf, age <=70), cor.test(age, friend_count, method = 'pearson'))

# Spearmen rho
c2 <- with(subset(pf, age <=70), cor.test(age, friend_count, method = 'spearman'))

# Kendall tau (NOTE: kendall take VERY LONG...minutes!!)
c3 <- with(subset(pf, age <=70), cor.test(age, friend_count, method = 'kendall'))

print('Pearson r: ', c1$estimate)
print('Spearman rho: ', c2$estimate)
print('Kendall tau: ', c3$estimate)
```

## Create Scatterplots
Notes:

```{r}
ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) +
  geom_point(alpha = 1/10) +
  coord_cartesian(xlim = c(0,1000), ylim = c(0,1000))
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) +
  geom_point(alpha = 1/10) +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) + 
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

data(Mitchell)
?Mitchell
head(Mitchell)

mf <- Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(data = mf, aes(x = Month, y = Temp)) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(mf$Month, mf$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = mf, aes(x = Month, y = Temp)) +
  geom_point() + 
  scale_x_continuous(breaks = seq(0,range(mf$Month)[2],12))
```

***

### A New Perspective

What do you notice?
Response:
the cyclical signal in the temperature...indicative of seasonal trends in the soild temperature.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months = pf$age + (1 - pf$dob_month/12)
print(pf$age_with_months)
```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>% 
  group_by(age_with_months) %>% 
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>% 
  arrange(age_with_months)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
pf.fc_by_age_months2 <- pf %>% 
  subset(age_with_months < 71) %>% 
  group_by(age_with_months) %>% 
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>% 
  arrange(age_with_months)

ggplot(data = pf.fc_by_age_months2, aes(x = age_with_months, y = friend_count_mean)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

