---
title: 'Lesson_6: Problem Set - Explore Two Variables'
author: "Greg Clunies"
date: "3/18/2018"
output: html_document
---


# setup notebook for analysis
```{r}
# install.packages("tidyverse")
library(tidyverse)  # dplyr, tidyr, ggplot2, etc.

setwd('/Users/GregoryClunies/Repos/Udacity/Exploratory_Data_Analysis/L6_ProbelmSet_ExploreTwoVariables')

data(diamonds)
names(diamonds)
?diamonds
```


# Q1 - Price vs x
```{r}
ggplot(data = subset(diamonds, diamonds$x != 0), aes(x = x, y = price)) +
  geom_point(alpha = 1/20)
```
## Q2 - Findings - price vs. x
There seems to be an exponential relationship between diamond length (x) and price. There appears to be some error values where the diamonds has no length. 

Plotting the natural logarithm of price vs x produces what appears to be a linear relationship, further suggesting that the relationship between x and price is exponential (see plot below).

```{r}
ggplot(data = subset(diamonds, diamonds$x != 0), aes(x = x, y = log(price))) +
  geom_point(alpha = 1/10)
```

## Q3 - Correlations
```{r}
cor.test(diamonds$x, diamonds$price)
cor.test(diamonds$y, diamonds$price)
cor.test(diamonds$z, diamonds$price)
```

##Q4 - price vs. depth
```{r}
ggplot(data = diamonds, aes(x = depth, y = price)) +
  geom_point(alpha = 1/20)
```

##Q5 - Adjustments - price vs. depth
```{r}
ggplot(data = diamonds, aes(x = depth, y = price)) +
  geom_point(alpha = 1/100) + 
  scale_x_continuous(breaks = seq(50, 70, 2), limits = c(58,66))
```
## Q6 Typical Depth Range
Most diamonds have depths ranging from **60** to **64**.

## Q7 Correlation - price and depth
Most diamonds have depths ranging from **60** to **64**.
```{r}
with(diamonds, cor.test(depth, price))
```
The correlation coeff indicates a non-linear relationship. Although the scatter data appears to trends straight up (generally), this would mean an infinite slope! We could suggest that based on the scatter and correlation coeff., r, that price is independent of diamond depth.

## Q8 - price vs. carat
```{r}
ggplot(data = diamonds, aes(x = carat, y = price)) + 
  geom_point(alpha = 1/20) +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```

## Q9 - price vs. volume
```{r}
diamonds$volume <- with(diamonds, x * y * z)

ggplot(data = diamonds, aes(x = volume, y = price)) + 
  geom_point()
```

## Q10 - Findings - price vs. volume
There are some extreme outliers:
- volumes of zero!
- volume near 4000!
- very cheap diamonds with volume of approx. 900

## Q11 Correlations on Subsets
```{r}
df <- subset(diamonds, diamonds$volume < 800 & diamonds$volume != 0)

cor.test(df$price, df$volume)
```

## Q12 - Adjustments - price vs. volume
```{r}
df <- subset(diamonds, diamonds$volume < 800 & diamonds$volume != 0)

ggplot(data = df, aes(x = volume, y = price)) +
  geom_point(alpha = 1/50) + 
  geom_smooth(method = 'lm')
```
The linear is a reasonable model for estimating price based on volume given the data. An r value of 0.92 indicates a pretty strong linear relationship. But we can see from the plot it does not capture all points, hence r < 1.

## Q13 - Mean Price by Clarity

diamonadsByClarity <- diamonds %>% 
```{r}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n())
```

## Q14 - Bar Charts of Mean Price
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(diamonds_mp_by_clarity, aes(x = clarity, y = mean_price)) +
  geom_col()

p2 <- ggplot(diamonds_mp_by_color, aes(x = color, y = mean_price)) +
  geom_col()

library(gridExtra)
grid.arrange(p1, p2)
```

## Q15 - Trends in Mean Price
Generally, the worse graded diamonds, by color and clarity, are more expensive. This seems counter intuitive...

## Q16
Gaminder data --- skipped this question as it donest add much additional value.