---
title: 'Lesson_4: Problem Set - Explore One Variable'
author: "Greg Clunies"
date: "3/18/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set Working Directory
```{r}
# macbook path
setwd('/Users/GregoryClunies/Repos/Udacity/Exploratory_Data_Analysis/L4_ProblemSet_ExploreOneVariable')

# windows path

```


## Load data from ggplot package
```{r}
library(ggplot2)  #must load the ggplot package first
data(diamonds)  #loads the diamonds data set since it comes with the ggplot package
```

## Q1
```{r}
summary(diamonds)
str(diamonds)
?diamonds
```

## Q2 - Histogram
```{r}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram()
```

## Q3 - Describe Shape, Summary Stats
```{r}
summary(diamonds$price)
```
Data is right skewed and long tailed. There is a slight spike around the \$4500 price. Prices range from \$326 to \$18823 (!!!). The median price is \$2401. The mean is \$3933.

## Q4 - Diamond Counts
```{r}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```

## Q5 - Cheaper Diamonds
```{r}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram(binwidth=50, color = 'black', fill = 'orange') +
  scale_x_continuous(limits = c(0,3000), breaks = seq(0,3000,250))

ggsave('CheapDiamonds_Histogram.png')
```

## Q6 - Histograms: Price by Cut
```{r}
ggplot(data = diamonds, aes(x = price)) +
  geom_histogram(binwidth = 50, color = 'black', fill = 'orange') +
  scale_x_continuous(breaks = seq(0,5000,500), limits = c(0, 5000)) +
  facet_wrap(~cut, ncol = 1)
```
It seems strange that so many "ideal" cut diamonds are available at such a 
low price. More investigation later?

## Q7 - Price by cut
```{r}
by(diamonds$price, diamonds$cut, summary)
```

## Q8 - Scales and Multiple Histograms
```{r}
# Run the code below in R Studio to generate
# the histogram as a reminder
qplot(x = price, data = diamonds) + 
  facet_wrap(~cut)

# 'free' the y-axis scales fro each facet
qplot(x = price, data = diamonds) + 
  facet_wrap(~cut, scales = 'free_y')
```

## Q9 - Price per Carat by Cut
```{r}
diamonds$price_per_carat <- diamonds$price/diamonds$carat

ggplot(data = diamonds, aes(x = price_per_carat)) +
  geom_histogram(binwidth = 0.05, color = 'black', fill = 'orange') +
  scale_x_log10() +
  facet_wrap(~cut, scales = 'free_y') +
  annotation_logticks(sides = 'b')
```

## Q10 - Price Box Plots
```{r}
by(diamonds$price , diamonds$cut, summary)

ggplot(data = diamonds, aes(x = cut, y = price)) + 
  geom_boxplot() + 
  coord_cartesian(ylim = c(0, 7500))

# save figure
ggsave('Price_by_Cut_Boxplot.png')
```
It's strange that the average price of diamonds trends downwards from Fair to Very Good cuts, then jumps to it's highest average for Premium cuts, but is LOWEST for premium cuts. Also, all cuts have extremely high maximum prices. These observations suggest that cut alone does not account for the variation in price. 

## Q11 Interquartile Range - IQR
```{r}
by(diamonds$price, diamonds$color, summary)

?diamonds

IQR_D = IQR(subset(diamonds, color == 'D')$price)
IQR_D

IQR_J = IQR(subset(diamonds, color == 'J')$price)
IQR_J
```

## Q12 - Price per Carat Box Plots by Color
```{r}
ggplot(data = diamonds, aes(x = color, y = price_per_carat)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0,6000))
```

## Q13 - Carat Frequency Polygon
```{r}
ggplot(data = diamonds, aes(x = carat)) +
  geom_freqpoly(binwidth = 0.1) +
  scale_x_continuous(breaks = seq(0,5,0.2), limits = c(0,5)) +
  scale_y_continuous(breaks = seq(0,12000,1000)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=8))
```

## Q14 - Gapminder Data
```{r}
library(tidyr)
library(dplyr)

df <- read.csv("indicator_extreme temp affected.csv", check.names = F, stringsAsFactors=FALSE)

colnames(df)[1] <- "country"

df <- df[-dim(df)[1],-dim(df)[2]]
```
This dataset summarizes the Total number of people getting affected, injured or killed in extreme temperature during the given year, by country.

The dataset in its current state is not in a very friendly format for plotting, so let's clean this up a bit using gather(). Remember, we want a dataframe where each variable has it's own column, and each observation has its own row.

```{r}
df <- gather(df, key = "year", value = "num_sick", 2:37)
```
This looks good, time for some plots.

Let's start by seeing which country has had the most temperature related sicknesses over the ENTIRE dataset duration.
```{r}
df2 <- df %>%
  group_by(country) %>% 
  summarise(total_sick = sum(num_sick))

### NEVER FINISHED - kind of bullshit lesson
```
